Jury members' comments on the private PhD defense of \linebreak Gustavo Quintana-Carapia

### 2020-08-21 internal defense feedback/Answers-jury-members-report_pending.tex
Second draft before public defense

Thank you to the jury members for reviewing the thesis manuscript and for giving suggestions for improvement. 
The manuscript changes due to the comments are highlighted in \color{blue} blue fontcolor\color{black}.

@@

\begin{itemize}
	\item The manuscript text should better position the PhD to related work. Improve the literature review and discuss (and cite) the state of the art.

	{\bfseries The introduction of the chapters now includes the state of the art to position the PhD work with respect to the literature.}

	\item  Add more clearly the motivation for the PhD research work by including motivational examples to the introduction.

	{\bfseries Motivational examples have been added to the Introduction. }

	\item  Add a discussion with comparison of existing EIV methods. Give, with the positive and negative aspects in mind, a better justification for why you use a (R)LS method, and not, e.g., (R)IV method.

	{\bfseries In Section 4.1.4, a discussion that compares IV methods with EIV methods has been added.}

	\item  Discuss more clearly the aspects of scalability of the methods and their limitations in terms of model order, size of data sets (e.g., related to the inversion of LS matrices).
	\item  The equations on pages 12 and 13 seemingly contain errors and should be corrected.

	{\bfseries The equations on pages 12 and 13 have been corrected.}

	\item  The figure shown on page 66 was made with a single experiment; it is desirable to see the (averaged) effect on several experiments.

	{\bfseries A series of 100 runs of the maximum-likelihod estimation method has been simulated and the results have been added to the manuscript in Section 6.3.2.}

@@

\begin{itemize}
	\item You develop results for uncertainty quantification; what type of uncertainties  can your approach deal with and which levels of noise can you consider? Elaborate on SNR levels; I’m also very interested in scalability, so how big the matrices can be.
	\item  Concerning the scalability to data: eqn 3.6: how scalable is the approach in all aspects (order, data, linked to inversion for LS matrix); can you deal with hundreds, thousands of data?

	{\bfseries The following paragraph has been added in Section 3.2 to discuss the scalability of the data-driven step input estimation method.}

	\color{blue}
    The computational complexity of the RLS algorithm solution to the system of equations (3.14) is $O \left( \left( n+1 \right)^2 \right)$.
    The largest computational requirement is for the initialization of the algorithm, where a matrix inverse is required when the number of samples is just enough to have a square matrix $\widetilde{\mathbf{K}}_{n+1}$.
    From there on, the RLS algorithm updates the solution with linear complexity.
    Therefore, even though a large number of samples $N$ makes the matrix in the system of equations (3.14) increase in the number of rows, the estimation of the solution does not increase in complexity. 
    The data-driven step input estimation method is then scalable for any sensor of order $n$, and can be executed in devices with limited computational resources, provided that the computation of the  $n+1 \times n+1$ inverse matrix is feasible.
    \color{black}

	\item  What recommendation up to scale/order can method work with. What are the limitations? It would be good to have more precise discussion about these aspects in dissertation: what do you mean by large scalability: 100, $10^4$, ...? A more detailed discussion is desirable.
	\item  On slide 25 (noisy graphs) it would be good to see a statistical analysis (more experiments than one).

    {\bfseries 100 runs of the maximum-likelihod estimation method have been simulated and their results have been added to the manuscript in Section 6.3.2.}

	\color{blue} 
    The maximum-likelihood (ML) method estimated the affine input parameters from the sensor model transient response.
    A total of 100 runs of the method were obtained, using different realizations of measurement noise.
    The measurement noise variance was set to have an SNR of 40 dB.
    The ML method used the first 50 samples to initialize the optimization variables and updated the variables every $N_{\mathrm{s}} = 5$ samples.
    Figure \ref{fig:rele_lo_40dB_s10} shows the average of the observed relative errors in the estimation of the parameters $\widehat{a}$, $\widehat{b}$, $\widehat{x}_{\mathrm{ini,1}}$, and $\widehat{x}_{\mathrm{ini,2}}$.

    \renewcommand{\thefigure}{6.8}
    \begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterRampInput/fig/Fig_7.pdf} 
    \caption{ \label{fig:rele_lo_40dB_s10} The affine input parameters and the sensor's initial conditions are estimated with the ML method. After three iterations, \color{blue} the average of the relative error is smaller than 5 \% for each and every of the estimates. The estimate $\widehat{x}_{\mathrm{ini,2}}$ has the larger relative error near to 1 \%. \color{black} }
    \end{figure}

    Figure \ref{fig:cov_lo_40dB_s1} shows \color{blue} the average \color{black} of the variances that were observed on the diagonal of the covariance matrix $\mathbf{J}$. 
    We can see that the estimation variances decrease as more samples are processed.
    Moreover, the estimation variances of $\widehat{a}$ and $\widehat{b}$ obtained with the ML method are lower than the corresponding estimation MSE errors obtained from a Monte Carlo simulation of the subspace method.

    \renewcommand{\thefigure}{6.9}
    \begin{figure}[!htbp]
    \centering
    \includegraphics[width=0.6\columnwidth]{../Gus-thesis/ChapterRampInput/fig/Fig_8.pdf} 
    \caption{ \label{fig:cov_lo_40dB_s1} The variances of the ML estimates are calculated using the information provided by the analytic Jacobian. \color{blue} The estimates $\widehat{x}_{\mathrm{ini,1}}$ and $\widehat{a}$ have the smaller and larger variances, respectively. \color{black}  }
    \end{figure}

    \color{black}

@@

\begin{itemize}
	\item The Taylor expansion is computed of order 2, what is the accuracy of expansion with respect to system, noise level,... Can you come up with a priori bound on error that will allow you to assess the quality of the expansion? Can you predict where the solution is going to be before computing Taylor expansion?

	{\bfseries An a priori bound on the error introduced by the Taylor series expansion can be obtained in terms of the spectral radius of the matrix $\mathbf{M}$. In Section 4.1 the following description has been added: }

    \color{blue} An a priori bound on the error introduced by the second order Taylor series expansion can be expressed in terms of the spectral radius $\rho(\mathbf{M})$.
    Considering that the Taylor series expansion is an infinite sum we have that
    \begin{equation} \tag{4.4} \begin{aligned} (\mathbf{I} + \mathbf{M})^{-1} &= \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 + \ldots \\
    &=  \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3 + \ldots \right) \\
    &=  \mathbf{I} - \mathbf{M} + \mathbf{M}^2 - \mathbf{M}^3  (\mathbf{I} + \mathbf{M})^{-1} \end{aligned} \end{equation} 
    Therefore we can obtain a bound for the second error expansion error if we take a matrix norm as follows
    \begin{equation} \tag{4.5} \begin{aligned} \| (\mathbf{I} + \mathbf{M})^{-1} - ( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 ) \| &=  \| \mathbf{M}^3  (\mathbf{I} + \mathbf{M})^{-1} \| \\
    & \leq \| \mathbf{M}^3 \| \| (\mathbf{I} + \mathbf{M})^{-1} \| \leq \dfrac{\| \mathbf{M} \|^3}{ 1 -  \| \mathbf{M} \|} = \dfrac{\rho(\mathbf{M})^3}{1 - \rho(\mathbf{M})}  \end{aligned} \end{equation} 
	\color{black}

	\item  Concerning other sorts of prior estimations, e.g., given by explicit function theorems. I wonder if you could get something and apply this bound in order to control the bound of the 2nd order expansion. There exist some good results that could help you devise something a priori which can control the error you are making. Neuberger's results (quite old now) but may be handy for this kind of problem. This can be investigated further.

	\color{blue}
	Neuberger's paper \cite{Neuberger07} introduces five theorems that are consequence of the continuous Newton method.
	The Newton method aims to find a zero of the function $F: R^n \rightarrow R^n$ by iterating
	\begin{equation} z_{k+1} = z_k - \left(F' (z_k) \right)^{-1} F(z_k) \end{equation}
	for $k=0,1,2,\ldots$, and for a given initial $z_0$, assuming that $\left(F' (y) \right)^{-1}$ exists for some $y \in R^n$.
	A domain of attraction is the set of all initial values $z_0$ that lead to a root of $F$ after the convergence $z_0, z_1, z_2,\ldots$.
	The damped Newton method  
	\begin{equation} z_{k+1} = z_k - \delta_k \left(F' (z_k) \right)^{-1} F(z_k) \end{equation}
	prevents convergence issues, such as chaotic domains of attraction, with $\delta_1, \delta_2, \ldots \in (0, 1)$. 
	The continuous Newton method is a sequence of damped Newton methods. To implement the continuous Newton method, select $T>0$ and perfom $m$ runs of the damped Newton method with $\delta_k = T/m$, for $k=1,2,\ldots, m$.
	The result $z_{k+1}$ is the zero estimation $x_m$, and if the sequence of estimates $x_1, x_2, \ldots$ converges, then it converges to the result of the continuous Newton method.

	The objective of the continuous Newton method is to find a function $z: [0, \infty) \rightarrow R^n$, so that
    \begin{equation} z(0) = x \in R^n, \quad z' (t) = - \left(F' (z(t)) \right)^{-1} F(z(t)), \quad t \geq 0, \label{eqn:contNewton}\end{equation}
    subject to the existence of the limit $u=\mathrm{lim}_{t \rightarrow \infty}{z(t)}$, that satisfies $F(u)=0$.
    This implies 
    \begin{equation}  F' (z(t)) z' (t) = - F(z(t)) , \end{equation}
    that is equivalent to
    \begin{equation}  \left(F \circ z\right)' (t) = - F(z(t)) , \end{equation}
    from where we get
    \begin{equation}  F(z(t)) = e^{-t} F(z(0)) . \end{equation}
    Thus, the residual $F(z(t))$ only changes in magnitude and not in direction.
    This property is essential for the results of the theorems in \cite{Neuberger07}. 
    In particular, Theorem 5 gives the conditions for the continuous Newton method to ensure finding a root of the function.

    \begin{customthm}{5} \label{thm:five}
    Suppose that each of $H$, $J$, and $K$ is a Banach space, with $H$ compactly embedded in $J$, that $r>0$, and that $G:B_r(0) \rightarrow K$ is continuous as a function on $J$. Suppose also that $g$ belongs to $K$ and that for each $y$ in $b_r(0)$ there is an $h$ in $B_r(0)$, where $b_r(0)$ and $B_r(0)$ are open and closed balls in $H$ of radius $r$ centered at 0, such that
    \begin{equation} \lim\limits_{t \to 0+} -{\dfrac{1}{t} \left(G(y+th) - G(y)\right) = g}. \end{equation}
    Then, there exists $u$ in $B_r(0)$ such that $G(u)=g$.
    \end{customthm}
    %$F:B_r(0) \rightarrow K$

    The theorem implies that $u$ is a zero of the function  defined as $F(y) = G(y) - g$.  
    Since $G$ is a continuous function and if $\left(G' (y) \right)^{-1}$ exists for each $y$ in $B_r(0)$, a metric for the validity of the Theorem (\ref{thm:five}) hypothesis is the inequality $\left\Vert  \left( G' (y) \right)^{-1} g \right\Vert \leq r$.

    To link the data-driven step-input estimation problem with this theorem, we can take the system of linear equations $\widetilde{\mathbf{y}} = \widetilde{\mathbf{K}} \bm{\theta}$, then define $G(y) = \widetilde{\mathbf{K}} \bm{\theta}$, and $g = \widetilde{\mathbf{y}}$, so that we have $F(\bm{\theta}) = G(\bm{\theta}) - g$. The derivative of the function $G(\bm{\theta})$ is given by $G'(\bm{\theta}) = \widetilde{\mathbf{K}} \mathbf{I}$. Using the pseudo-inverse of $\widetilde{\mathbf{K}}$, the metric $\left\Vert  \widetilde{\mathbf{K}}^{\dagger} \widetilde{\mathbf{y}} \right\Vert \leq r$ indicates that the estimate $\hat{\bm{\theta}}$ is found inside a ball of radius $r$.
    Thus,  
	\color{black}

@@

\begin{itemize}
	\item In the first or second sentence: “users cannot wait for a steady state regime” . Do you have specific examples of real industrial systems for which it is necessary to have continuous read-out before steady-state is reached? - Further on the industrial part and selling sensors with data-driven parts. Can you explain to experts that your solution is better? Sensors will be used for ages; why not spend one day to do system identification and get a reliable model to make a compensator, instead of using your approach?

	{\bfseries The fast weighing of parcels in the post office is a typical example of the real life systems where the time is limited to perform the measurements. Another one that we see nowadays is the temperature measurement of people in airports or train stations to prevent the spread of corona virus. The data driven method can be implemented in portable devices with limited computational resources. In industrial plants were a time slot can be devoted to identify the measurement system and implement a compensation method, the solution might be useful only for a certain amount of time. It is assumed that the model approximation deviates due to changes in the system elements, leading to inaccuracies in the inpiut estimation. There is a need for adaptive methods to avoid stopping the production lines to repeat the system identification procedure. The data-driven method is such an adaptive method. }

	\item  Chapter 2, Section 2.2.1. Can you give me some classic solutions/algorithms instead of LS in general? (Say I'm not familiar with EIV or TLS?) Usually with LS problems: multiply with $K^\top$ and compute pseudoinverse. Do you know another solution where they don't use K on the left-hand-side but something else? To get rid of noise by using a matrix on the left-hand-side, is related to EIV. There are some recursive solutions for this. It could be interesting to compare with such IV solutions.
	\item  In Chapter 3: at beginning you say to use linear systems theory to model a sensor. Can you discuss this assumption: is it true in practice? Can you guarantee it behaves like an LTI system?
	\item  Equation 3.1: an LTI system, described with SS representation. You only put output noise on this representation. People sometimes also add process noise. Why? How to modify your approach if there is process noise? Is it still white in this case?
	\item  On page 12 standard approach top of page equation: there are some mistakes in the equations! there is no epsilon (noise) in the equation. But then you do an SVD of H and estimate from there. How does it work if you have noise? Have you tried implementing it and see what you get?

	{\bfseries The step input estimation based on singular value decomposition works only with exact data. When  noise perturbes the sensor response the method cannot guarantee the step input estimation. The singular value decomposition of a Hankel matrix constructed from a perturbed sensor response provides only an approximation of the singular values and the singular vectors of this Hankel matrix. Then, the step input estimation obtained with this method from perturbed sensor response is only a rough estimation.     }

	\item  On page 12: I’m not sure the equations at the end of the page are correct, e.g., Y=G ubar ... not sure there is x; same for rest; There seem to be many mistakes here and next pages. Be careful with the equations!

	{\bfseries The equations have been reviewed and the typos were corrected.}

	\item  On page 13, you assume persistency of excitation of order $L$. How do you check this? In your case the input is constant; how to guarantee that the assumption is met?

	{\bfseries The persistency of excitation of a signal is verified by looking at the rank of the Hankel matrix constructed from the consecutive samples of this signal. When the Hankel matrix of $L+1$ block rows has a rank lower or equal than L, then we say that the signal is persistently exciting of order $L$. A step input has a very low order of persistency $L=1$, but the step respose of a system has a larger order of persistency. In fact, we assume that if the persistency of excitation of the step response is sufficiently high, then the columns of the Hankel matrix constructed from the step response span a linear space with all the responses of the equivalent augmented autonomous system where the step input is considered an augmented initial condition.}

	\item  In Chapter 4, p 17, Section 4.1.1: You assume that perturbations are IID and must be normally distributed. Why using an assumption of normally distributed? I would think that the kind of distribution is important for CRLB, but I’m not sure you need it for eqns 4.8, 4.9, 4.10,... IID is important, normality perhaps not.

	{\bfseries The following paragraph has been added in Section 4.1 to discuss on the need for the normality assumption.  }

	\color{blue}
	The normality assumption is necessary to provide more prior information into the method formulation.
    The second order Taylor series expansion (4.4) is developed considering the perturbation noise that gets into the regression matrix from the sensor transient response.
    Assuming only that the perturbation noise is distributed with zero mean and variance $\sigma_{\bm{\epsilon}}^2$ is not enough.
    If we additionally assume normality for the perturbation noise, then we benefit from the knowledge of the third moment equal to zero due to symmetry, and the fourth moment equal to three times the squared variance\color{black}.

	\item   In equation 4.11 you have $\sigma_eps^2 + \sigma_e^2$; how to get in practice $\sigma_e$? Do you have to do this each time? Recursively?

	{\bfseries The measurementy noise variance is prior information and in Chapter 5 we describe two ways of estimating $\sigma_{\bm{\epsilon}}$ In summary, the measurement noise can be extracted fom the sensor ressponse once the transient effects are smaller than the noise floor, or from the sensor response power spectrum. The online estimation of the measurement noise, simultaneously with the online step input estimation is a topic for future research.   }

	\item   Switch to 4.2, simulation results: first part dedicated to K, randomly generated; not sure if still possible, but in all your results it would be nice to have somewhere explicitly K = ... (even if randomly generated). You use only one realization also for matrices A, B, C, D in the text, so it can be added explicitly.
	\item   in Fig 4.1 the bottom figure also discusses the structured case. Why do you talk about structured results in a section about the unstructured case?

	{\bfseries Section 4.2 has been improved to describe the result next to its discussion.  }

	\item   In 4.1 uses a constant K, but in 4.2.2 you use A, B, C, D, without any links to K. You compare the two curves generated from different data sets… It is strange to compare them. I don't see the link between both.

	{\bfseries Section 4.2 has been reestructured and the simulation for the unstructured case were redone to establish a link between the two cases. The exact data is the same in the two cases and the perturbation noise follows the corresponding rules of each case, with the same variance noise levels. In this way the comparison between the two cases makes more sense.  }

@@

	\item   I was surprised to read in Chapter 5 something very close to what I read in chapter 4. In both cases generated A, B, C. In 5.1 again oscillating 5th order system, similar conclusions, ... What is the added value of 5.1 wrt 4.2.2?

	{\bfseries The derivation described in Chapter 4 was conducted with a second order mass-spring-damper weighing system example on mind, and the results observed were submitted in the corresponding paper. Afterwards, when we moved to the practical implementation, we faced the increase of the model order after setting up the weighing experiment. The lowest order we can assume is fifth order. So we decided to repeat the simulations and these new results serve two purposes: a transition from simulation to experimental results, and also to get an insight on the impact of the bias and variance predictions after increasing the model order.    }

	\item   On the impact of selected order on results. You were talking about a related bias/variance trade off. So, do you know in the literature some solutions to avoid overfitting? Can you cite something? Can you use something in your case to test for a good order of system?
	\item   Do you know of cross-validation: split data into two parts or several parts; some for estimation, some for validation? Have you tried this? Can this be used here to select good orders? Better results? Influence of system order on bias/variance; if I only use one dataset it is complicated and usually you can increase complexity to mimic a data set, but to avoid overfitting (too high order) use a test/validation data set. with this i should see what is the best value... Have you tried this?
	\item   In Section 6.2.2. Why do you call it a max likelihood approach? To me it is not one. Why do you call it like this? Where is the likelihood? Depending on which likelihood (eg. Gaussian, ...) it leads to LS. Don't' call it ML if you dont give distributions on ... In this case it is better to call it a nonlinear minimization procedure. Not clear why you call it ML.

	{\bfseries The method description in Section 6.2.2 has been modified to explain why the method is a maximum likelihood method.  }

@@

\begin{itemize}
	\item When solving algorithms. Do you opt for closed-form solutions, or do you use optimization based algorithms? How to solve LS problems? Do you solve them using pseudo-inverse or do you do gradient descent?
	\item  Today there is an overwhelming amount of data sensed and communicated; How would you apply your method in a scenario where dimensionality increases significantly? This relates to the questions of previous jury members; what is the implication of using numerical versus closed form solutions?
	\item  What if the $K^\top K$ matrix is not invertible? What if you have fewer samples than unknowns; what can you do? Solutions that assume structure like in compressed sensing, regularization,... can you use them, what would be the impact of that?
	\item  You mention that you improve complexity with your approach, but I missed a complexity analysis. Did you do this? Anything in processing time, flops,...?
	\item  You mentioned deep learning could be applied; could you show how? An issue is that neural nets are trained in high SNR data, but applied in engineering with lower SNR. This clogs performance; any ideas? Is this a limitation?
	\item  You mention that you consider Kalman filters; have you seen approaches where they write Deep Learning in form of Kalman Filter; Could this be applicable?
	\item  Recommendation: if you revise the thesis: maybe change "I propose" to "we propose".

	{\bfseries The formulation “I propose” was replaced by “we propose”. }

@@

\begin{itemize}
	\item Your main assumption is that a sensor outputs a steady state value that is proportional to the to-be-measured quantity. This means that you assume that the sensor is a linear system. What if this assumption is not met, and the sensor has a saturation for high values? Can you still use your methods?
	\item  Sensors are physical devices and deliver a continuous-time output. The models that you use are in discrete-time. Are there any risks that you miss information by not sampling fast enough?

	{\bfseries In Chapter 3 has been added a discussion on the step invariance transformation. This defines the conditions of the input for describing exactly the system dynamics in discrete time after sampling the continuous system response. If the input is stepwise continuous, then the zero order hold sampling of the step response leads to an exact discrete time representation of the system. The method under study in this thesis was developed for step input, then we can rely on the discrete time representation. There exist corresponding invariant transformations for piecewise linear inputs and for piecewise polynomial inputs that lead to exact representations from first and second order sampling of the system response, respectively.  }

@@

\section*{Roger Vounckx}

\begin{itemize}
	\item Two practical questions. On page 35-49 you consider a weight of 138.x grams. You sample relatively slowly (4 kHz) and after 500 samples we can consider that we have correct estimation. If I have a container that fills up and weight is constantly changing. Can we tackle this?

	{\bfseries Yes, that is exactly the idea considered by the affine input estimation method. Since the data-driven estimation method can track the changes of the input, the parameters of an input model that describes the constant variation can be recovered if we add the forgetting factor to the recursive solution algorithm. }

	\item  Consider an optical fiber link with a detector capable of 100 million measurements per second. Can I speed up the link by using your method? You have to do calculations: how high can you go in speed before running into troubles with DSP?

	{\bfseries The optical fiber link transmits digital pulses. Since the period would be 1 ns, the step input method implementation would require to sample in much shorter time. Assuming that the duty cyce is 50 \%, and that the number of samples required for the step detection are 200, then the sampling time should be 1.25 fs or lower. This could be a constraint for the commutation circuits and we would be reaching a feasibility limit. However, theoretically the method would perform detecting the pulses level changes.  }

@@

\begin{itemize}
	\item Returning to a question by Guillaume Mercere: on page 17; do you need normality? You made 2nd order taylor series expansion. Which moment do you need to estimate? Look at the equations page 16. There are 3rd, 4th order moments,... you need to know something more about the noise; this is one of the reasons to make Gaussian assumption: 3rd order moment is 0 (symmetry); 4th order for Gaussian is easy to have an explicit relationship.

    {\bfseries Section 4.1 has been appended with a paragraph to discuss on the need for the normality assumption.  }

@@

\section*{Ivan Markovsky}

\begin{itemize}
	\item About scalability of algorithms: You said in the very beginning that it can be very fast compared with system identification. This is the point: can you think of a good answer for these questions on scalability? Nikos Deligiannis mentioned computational analysis. You can directly say for (R)LS what is the computational complexity. Order three if full matrix inversion, now per iteration only linear. So you can scale to very large data sizes.

	{\bfseries The last paragraphs of Section 3.2 have been included to discuss the scalability of the data-driven step input estimation method.  }

@@

\begin{itemize}
	\item It is important for me to know why these new developments are generated. That is the reason why I suggest adding sections or paragraphs dedicated to
    \begin{itemize}
        \item the reasons why, in the industry, we need such a new technique. As I said during the defense, do you have a list of sensors for which waiting for the steady regime for getting the measurement is not possible? Why do you need a fast algorithm? What are the reasons why industry people need fast algorithms? In the end, it seems that, for you, it is more important to have a fast biased result than an unbiased estimate. Why? What do you mean by fast?
        \item  the literature on EIV estimation problems. As shortly mentioned in Chapter 2, there are solutions in the literature (by the way, you never mention the recent book written by Torsten Soderstrom in your document which is strange for me). But you do not describe them (even shortly) and, more importantly, you do not criticize them. Not in general but, at least, explain why you do not use them, what are their advantages and drawbacks and, finally, w.r.t. this list of positive and negative points, why you choose a (R)LS solution. E.g., why not using a recursive instrumental variable solution instead?

        {\bfseries A description of IV methods has been added in Section 4.1.4}

        \color{blue}
    \subsection*{4.1.4. Instrumental variables formulation of the data-driven step input estimation method}

    The errors-in-variables problem formulation (3.18) of the data-driven step input estimation method can be converted into an instrumental-variables problem.
    A guide to do this conversion is found in \cite{Soderstrom18} and in \cite{Pintelon12Book}. 
    The instrumental variable $\widetilde{\mathbf{Z}}$ has to be selected so that the normal equations (3.19) are expresed instead as
    \begin{equation} \tag{4.22} \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{K}} \bm{\theta} = \widetilde{\mathbf{Z}}^\top \widetilde{\mathbf{y}} . \label{eqn:neq_siv} \end{equation}
    The selection of the instrumental variable has to be done in such a way that is correlated with the regression matrix $\widetilde{\mathbf{K}}$, and uncorrelated with the perturbations $\bm{\epsilon}$.
    Assuming that the block-Hankel matrix $\widetilde{\mathbf{K}}$ is given by
    \begin{equation} \tag{4.23} \widetilde{\mathbf{K}} = \begin{bmatrix} G & \Delta \widetilde{y}(1) & \Delta \widetilde{y}(2) & \cdots & \Delta \widetilde{y}(n) \\ G & \Delta \widetilde{y}(2) & \Delta \widetilde{y}(3) & \cdots & \Delta \widetilde{y}(n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(n+1) & \Delta \widetilde{y}(n+2) & \cdots & \Delta \widetilde{y}(2n) \end{bmatrix} , \label{eqn:matrixK_r} \end{equation}
    then, the instrumental variable $\widetilde{\mathbf{Z}}$ should not contain the elements that construct $\widetilde{\mathbf{K}}$, i.e., the elements of the vector $\left( \widetilde{y}(1), \ldots, \widetilde{y}(2n+1) \right)^\top$.
    A possibility to generalize this idea is expressing the instrumental variable as
    \begin{equation} \tag{4.24} \widetilde{\mathbf{Z}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-4n) & \Delta \widetilde{y}(t-4n+1) & \cdots & \Delta \widetilde{y}(t-3n-1) \\ G & \Delta \widetilde{y}(t-4n+1) & \Delta \widetilde{y}(t-4n+2) & \cdots & \Delta \widetilde{y}(t-3n) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-3n) & \Delta \widetilde{y}(t-3n+1) & \cdots & \Delta \widetilde{y}(t-2n-1) \end{bmatrix} , \label{eqn:matrixZ_t} \end{equation}
    for a regression matrix
    \begin{equation} \tag{4.25} \widetilde{\mathbf{K}}(t) = \begin{bmatrix} G & \Delta \widetilde{y}(t-2n+1) & \Delta \widetilde{y}(t-2n+2) & \cdots & \Delta \widetilde{y}(t-n) \\ G & \Delta \widetilde{y}(t-2n+2) & \Delta \widetilde{y}(t-2n+3) & \cdots & \Delta \widetilde{y}(t-n+1) \\ \vdots & \vdots & \vdots & & \vdots \\ G & \Delta \widetilde{y}(t-n+1) & \Delta \widetilde{y}(t-n+2) & \cdots & \Delta \widetilde{y}(t) \end{bmatrix} , \label{eqn:matrixK_t} \end{equation}
    for $t = 4n+1, \ldots, N$.

    The difference with the EIV formulation is that this implementation of the IV method requires to wait more, until $4n+1$ samples of the sensor response are read, instead of $2n+1$ samples. 
    For sensors of small order this twice-the-time waiting does not represent a big change, but as the sensor order increases, the need for waiting can turn the method into impractical since a lot of time should pass before starting to estimate the input.

    There exists a generalized instrumental variables method (GIVE) that was proposed to identify dynamic models from input/output data. 
    GIVE is a bias-eliminating method that uses a generalized instrumental vector constructed from delayed values of the input and output signals \cite{Soderstrom18}.
    Since the aim of this thesis is to analize a method that directly estimates the input from the observed sensor output, without identifying explicitly a dynamic model of the sensor, the GIVE method is out of the scope of this work. 
    A future research can include studying the formulation of a data-driven input estimation method in the spirit of the GIVE method. 
    Nevertheless, a brief description of the GIVE method and its recursive version is given in the following paragraphs.

>>>>

@@

\begin{itemize}

	\item  Discuss more clearly the aspects of scalability of the methods and their limitations in terms of model order, size of data sets (e.g., related to the inversion of LS matrices).

@@

\begin{itemize}
	\item You develop results for uncertainty quantification; what type of uncertainties  can your approach deal with and which levels of noise can you consider? Elaborate on SNR levels; I’m also very interested in scalability, so how big the matrices can be.

	\item  What recommendation up to scale/order can method work with. What are the limitations? It would be good to have more precise discussion about these aspects in dissertation: what do you mean by large scalability: 100, $10^4$, ...? A more detailed discussion is desirable.

@@

\begin{itemize}

@@

\begin{itemize}

	\item  Chapter 2, Section 2.2.1. Can you give me some classic solutions/algorithms instead of LS in general? (Say I'm not familiar with EIV or TLS?) Usually with LS problems: multiply with $K^\top$ and compute pseudoinverse. Do you know another solution where they don't use K on the left-hand-side but something else? To get rid of noise by using a matrix on the left-hand-side, is related to EIV. There are some recursive solutions for this. It could be interesting to compare with such IV solutions.
	\item  In Chapter 3: at beginning you say to use linear systems theory to model a sensor. Can you discuss this assumption: is it true in practice? Can you guarantee it behaves like an LTI system?
	\item  Equation 3.1: an LTI system, described with SS representation. You only put output noise on this representation. People sometimes also add process noise. Why? How to modify your approach if there is process noise? Is it still white in this case?

	\item   Switch to 4.2, simulation results: first part dedicated to K, randomly generated; not sure if still possible, but in all your results it would be nice to have somewhere explicitly K = ... (even if randomly generated). You use only one realization also for matrices A, B, C, D in the text, so it can be added explicitly.

@@

	\item   On the impact of selected order on results. You were talking about a related bias/variance trade off. So, do you know in the literature some solutions to avoid overfitting? Can you cite something? Can you use something in your case to test for a good order of system?
	\item   Do you know of cross-validation: split data into two parts or several parts; some for estimation, some for validation? Have you tried this? Can this be used here to select good orders? Better results? Influence of system order on bias/variance; if I only use one dataset it is complicated and usually you can increase complexity to mimic a data set, but to avoid overfitting (too high order) use a test/validation data set. with this i should see what is the best value... Have you tried this?

@@

\begin{itemize}
	\item When solving algorithms. Do you opt for closed-form solutions, or do you use optimization based algorithms? How to solve LS problems? Do you solve them using pseudo-inverse or do you do gradient descent?
	\item  Today there is an overwhelming amount of data sensed and communicated; How would you apply your method in a scenario where dimensionality increases significantly? This relates to the questions of previous jury members; what is the implication of using numerical versus closed form solutions?
	\item  What if the $K^\top K$ matrix is not invertible? What if you have fewer samples than unknowns; what can you do? Solutions that assume structure like in compressed sensing, regularization,... can you use them, what would be the impact of that?
	\item  You mention that you improve complexity with your approach, but I missed a complexity analysis. Did you do this? Anything in processing time, flops,...?
	\item  You mentioned deep learning could be applied; could you show how? An issue is that neural nets are trained in high SNR data, but applied in engineering with lower SNR. This clogs performance; any ideas? Is this a limitation?
	\item  You mention that you consider Kalman filters; have you seen approaches where they write Deep Learning in form of Kalman Filter; Could this be applicable?

@@

\begin{itemize}
	\item Your main assumption is that a sensor outputs a steady state value that is proportional to the to-be-measured quantity. This means that you assume that the sensor is a linear system. What if this assumption is not met, and the sensor has a saturation for high values? Can you still use your methods?

@@

@@

\begin{itemize}

@@

@@

\begin{itemize}
	\item It is important for me to know why these new developments are generated. That is the reason why I suggest adding sections or paragraphs dedicated to
    \begin{itemize}
        \item the reasons why, in the industry, we need such a new technique. As I said during the defense, do you have a list of sensors for which waiting for the steady regime for getting the measurement is not possible? Why do you need a fast algorithm? What are the reasons why industry people need fast algorithms? In the end, it seems that, for you, it is more important to have a fast biased result than an unbiased estimate. Why? What do you mean by fast?

    \item  When you generate different simulation examples, explain why the new ones are interesting or, more importantly, what they bring w.r.t. the former simulations results. This is the case for instance when I compare Section 4.2.2 and Section 5.1.
